A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-07-20 Analysis/models.Rmd
add spr models, move floats

#data=readRDS("spr_pre_brms.rds")
data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
 priors <- c(
      set_prior("normal(0, 5)", class="b"),
      set_prior("exponential(.1)", class="sd", group="subject", coef="Intercept"),
      set_prior("exponential(.1)", class="sd", group="Word_ID", coef="Intercept"),
      set_prior("exponential(1)", class="sd"),
      set_prior("lkj(2)",       class="cor"),
      set_prior("exponential(.1)", class="sigma")
      )
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec) %>% 
    mutate(center_rt=scale(rt, center=T, scale=F))
  m1 <- brm(center_rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
              (surp_center + length_center + freq_center|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

@@

```{r diagnose-unhappiness}

sad <- read_rds(here("Analysis/models/gpt_spr.rds"))

@@

d <- data %>% filter(model=="GPT-2") %>% select(-rt, -Word_ID, -model, -subject)

cor(d) %>%as.tibble() %>%  mutate(across(where(is.numeric), ~round(.x,2))) %>% View()
```

>>>>

#data=readRDS("spr_pre_brms.rds")
data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
 priors <- c(
      set_prior("normal(350,50)", class="Intercept"),
      set_prior("normal(0, 5)", class="b"),
      set_prior("exponential(.01)", class="sd", group="subject", coef="Intercept"),
      set_prior("exponential(.01)", class="sd", group="Word_ID", coef="Intercept"),
     # set_prior("exponential(1)", class="sd"),
     # set_prior("lkj(2)",       class="cor"),
      set_prior("exponential(.01)", class="sigma")
      )
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec)
  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
              (1|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

@@

```{r diagnose-unhappiness, eval=F}

sad <- read_rds(here("Analysis/models/txl_spr.rds"))

@@

```
```{r summary-spr}

show_summary_spr <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

```{r process-spr-models, eval=F}
ngram_spr <- read_rds(here("Analysis/models/5-gram_spr.rds")) %>% show_summary_spr() %>% mutate(model="5-gram")
txl_spr <- read_rds(here("Analysis/models/txl_spr.rds")) %>% show_summary_spr() %>% mutate(model="TXL")
grnn_spr <- read_rds(here("Analysis/models/grnn_spr.rds")) %>% show_summary_spr() %>% mutate(model="GRNN")
gpt_spr <- read_rds(here("Analysis/models/gpt_spr.rds")) %>% show_summary_spr() %>% mutate(model="GPT-2")

summ_spr <- ngram_spr %>% union(txl_spr) %>% union(grnn_spr) %>% union(gpt_spr) %>% write_rds(here("Analysis/models/brms_spr_summ.rds"))

### 2021-07-20 Papers/Paper/manuscript.Rmd
add spr models, move floats

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "2"
    institution   : "Massachusetts Institute of Technology"
header-includes:
 - \usepackage{setspace}\singlespacing

@@

```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:diagram-cap)"}
knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
```
One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error. Thus, only participants who have processed the sentence and are paying attention contribute data at critical regions later in the sentence. However, this means we don't have data after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue. However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran. They recruited 181 participants, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020]. For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
TODO differences for linear model!

@@

```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:error-cap)"}
error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
  geom_point(size=1)+
  labs(x="Fraction words selected correctly",
       y="Mean Reaction Time (ms)")+
  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600), expand=F)+
    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
  guides(color="none")

@@

```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:gam-cap)"}

@@

knitr::kable(summ, format="latex", position="t",caption="Predictions from fitted Bayesian regression models. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
```

@@

read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on Maze data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests are reported. We also report log likelihoods of models with only one surprisal source.")

@@

```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos='t', fig.cap="(ref:spr-gam-cap)"}
gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

@@

```{r spr-table, eval=F}

all_tidy <- read_rds(here("Analysis/SPR_lmer_summ.rds")) %>% 
  select(Term=term, estimate, std.error, type) %>% 
  mutate(Ci_high=estimate+1.96*std.error,
         Ci_low=estimate-1.96*std.error,
         Range=str_c(round(estimate,1)," [",round(Ci_low, 1),", ",round(Ci_high,1),"]")) %>% 
  select(Term, type, Range) %>% 
  pivot_wider(id_cols=Term, names_from=type, values_from=Range) %>% 
    mutate(Term=factor(Term, 
                        levels=c("(Intercept)", 
                                 "surp_center",
                                 "length_center",
                                 "freq_center",
                                 "freq_center:length_center",
                                 "past_c_surp",
                                 "past_c_length",
                                 "past_c_freq",
                                 "past_c_freq:past_c_length")
                                 )) %>% 
  arrange(Term) %>% 
  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
                   "Freq x Length", "Past Surprisal",
                  "Past Length", "Past Freq", "Past Freq x Length"))

knitr::kable(all_tidy, format="latex", position="t",caption="Predictions from fitted linear regression models. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
```

As a comparison, we ran the same types of models on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020. As shown in Figure \@ref(fig:spr-gam), there is a roughly linear but fairly flat relationship between RT and surprisal. Note that the y-axis is fairly narrow, and so the predicted effect of changes in surprisal is fairly small. This is confirmed by linear models (see Table \@ref(tab:spr-table)); note that due to the size of the data, these models were fit differently than the A-maze models due to computational constraints. Surprisal and length effects are evident for both the current and past word, but we do not see frequency effects. The effects are much smaller than for A-maze, although some of this may be due to noise from spillover from prior words or differences in model structure.

Lastly, we are interested to know which of the models is better at predicting human behavior data, and whether they are complementary. I fit a series of models including the frequency and length predictors, and some number of past or present surprisal sources. This allows us to tell whether different models of surprisal are "picking up" on different things that affect reading time.

Similarly to the model comparison for Maze, we also conduct a model comparison on the SPR data shown in Table \@ref(tab:spr-compare). In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. Overall, F-values here are much lower than for Maze. TODO how to interpret!

@@

read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on SPR data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests are reported. We also report log likelihoods of models with only one surprisal source.")

>>>>

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "2"
    institution   : "Massachusetts Institute of Technology"
header-includes:
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.05}
 - \renewcommand{\topfraction}{0.8}
 - \renewcommand{\bottomfraction}{0.8}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}

@@

```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
```
One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error. Thus, only participants who have processed the sentence and are paying attention contribute data at critical regions later in the sentence. However, this means we don't have data after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue. However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran. They recruited 181 participants, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020]. For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
For linear models, we centered all predictors. We were unable to fit the full mixed effects model, so we used the best model that reasonably converged which had a by-subject random intercept and a per-word-token random intercept. We used regularizing priors(normal(350,50) for intercept, normal(0,5) for beta, exponential(.01) for sd and sigma). Models were run in brm [@burknerAdvancedBayesianMultilevel2018].

@@

```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:error-cap)"}
error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
  geom_point(size=1)+
  labs(x="Fraction words selected correctly",
       y="Mean Reaction Time (ms)")+
  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600), expand=F)+
    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
  guides(color="none")

@@

```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}

@@

knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
```

@@

read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests are reported. We also report log likelihoods of models with only one surprisal source.")

@@

```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

@@

```{r spr-table}

summ_spr <- read_rds(here("Analysis/models/brms_spr_summ.rds")) %>% 
  pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
  mutate(Term=factor(Term, 
                        levels=c("Intercept", 
                                 "surp_center",
                                 "length_center",
                                 "freq_center",
                                 "surp_center:length_center",
                                 "length_center:freq_center",
                                 "past_surp_center",
                                 "past_length_center",
                                 "past_freq_center",
                                 "past_surp_center:past_length_center",
                                 "past_length_center:past_freq_center",
                                 "past2_surp_center",
                                 "past2_length_center",
                                 "past2_freq_center",
                                 "past2_surp_center:past2_length_center",
                                 "past2_length_center:past2_freq_center",
                                 "past3_surp_center",
                                 "past3_length_center",
                                 "past3_freq_center",
                                 "past3_surp_center:past3_length_center",
                                 "past3_length_center:past3_freq_center"
                                 )
                                 )) %>% 
  arrange(Term) %>% 
  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency","Surp x Length", "Freq x Length",
                "Past Surprisal", "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length",
                "2Past Surprisal", "2Past Length", "2Past Freq", "2Past Surp x Length", "2Past Freq x Length",
                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length"))

knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models for SPR data. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")

As a comparison, we ran the same types of models on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020. As shown in Figure \@ref(fig:spr-gam), there is a roughly linear but fairly flat relationship between RT and surprisal. Note that the y-axis is fairly narrow, and so the predicted effect of changes in surprisal is fairly small. This is confirmed by linear models (see Table \@ref(tab:spr-table)). Surprisal and length effects are evident for the current word, and most models shows surprisal effects from the past word, but we do not see frequency effects. The effects are much smaller than for A-maze, even though this model accounts for spillover effects from more previous words.

Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare). In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. Overall, F-values here are much lower than for Maze.

@@

read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests are reported. We also report log likelihoods of models with only one surprisal source.")
